### 2018-04-27 manuscript/Manuscript.Rmd
More writing

---
title: "Do researchers preferentially collaborate with same-gendered colleagues?"
author:
- Luke Holman* and Claire Morandin
- '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
fontsize: 12pt
header-includes: 
  \usepackage[all]{hypcap} 
  \usepackage{amsmath} 
  \usepackage{booktabs}
  \usepackage{caption} 
  \usepackage[labelfont=bf]{caption} 
  \usepackage{titling} 
  \pretitle{\begin{flushleft}\LARGE}
  \posttitle{\par\end{flushleft}\vskip 0.5em} 
  \preauthor{\begin{flushleft}\large \lineskip 0.5em} 
  \postauthor{\par\end{flushleft}} 
  \predate{\begin{flushleft}\large} 
  \postdate{\par\end{flushleft}}
  \usepackage{fancyhdr} 
  \pagestyle{fancy} 
  \fancyhead[LO,LE]{\textsl{\leftmark}} 
  \rhead[]{Conference gender ratios} 
  \usepackage[noabbrev,capitalise]{cleveref}
  \usepackage{titlefoot} 
  \usepackage{amssymb} 
  \usepackage{rotating}
csl: the-american-naturalist.csl
abstract: Evidence suggests that women in academia are hindered by conscious and unconscious biases, and that many female researchers feel excluded from formal and informal opportunities for research collaboration. In addition to ensuring fairness and helping redress gender imbalance in the academic workforce, increasing women’s access to collaboration could help scientific progress by drawing on more of the available human capital. Here, we test whether researchers preferentially collaborate with same-gendered colleagues, using more stringent methods and a larger dataset than in past work. Our results reaffirm that researchers co-publish with colleagues of the same gender, and show that this ‘gender homophily’ is slightly stronger today than it was 10 years ago. Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority. Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research. \vspace{5mm} \par\noindent \textbf{Keywords:} Gender bias, Homophily, Scientific collaboration, Text mining, Women in STEM.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load some little files with numbers to be inserted into the MS
sample.size <- read.csv("sample.size.table.csv", stringsAsFactors = FALSE)
sample.size[,2] <- prettyNum(sample.size[,2], big.mark = ",")
sample.size[,3] <- prettyNum(sample.size[,3], big.mark = ",")
names(sample.size)[2:3] <- c("n2005", "n2015")

fig2 <- readRDS("fig2_numbers.RData") # For figure 2
n.journals <- fig2[1]
n.sig.homo <- fig2[2]
n.sig.hetero <- fig2[3]

R2 <- readRDS("R2_for_figure4.RData") # For Fig 4
```

\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia}
\newpage

# Introduction
Women are substantially underrepresented in the workforce in many branches of science, technology, engineering, mathematics, and medicine (STEMM), and they face additional challenges and inequities relative to men [e.g. @Lariviere_2013; @West_2013; @Elsevier_report; @Holman_2018]. On average, women occupy more junior positions (Wutte 2007, Reuben, Sapienza & Zingales, 2014) with lower salaries (Trower and Chait 2002, Umback 2007), receive less grant money (Hosek et al. 2005, OER 2005), are promoted more slowly (Hopkins et al. 2013; Long et al. 1993, Rosenfeld 1991, Zuckerman 1987), and are allocated fewer resources (European Commission 2009) and less research funding (Feldt 1986, Stack 2004, Larivière et al. 2011) than men. Experimental studies have demonstrated that researchers regard women's achievements less favourably than identical achievements by men (Moss et al. 2012; Knobloch et al. 2013). Multiple studies have found that women publish fewer papers per year than men (Cole and Zuckerman 1984, Long 1992, Larivière et al. 2013, Astronomy one in Nature) [check whether they control for career stage].

Because publishing, networking and collaboration are all instrumental to academic career advancement (Lee and Boheman 2005, Hopkins et al. 2012), numerous studies have tested for gender differences in these areas (reviewed in Table S1 in Holman et al. 2018; see also XXX). For example, studies have shown that women tend to be less involved in international collaboration than men (Lemoine 1995; Lewison 2001; Webster 2001; Larivière et al. 2011, Bozeman and Corley 2004), collaborate less with researchers from their own university department (Websters 2001), and have less prestigious collaborations compared to men (Long 1990), but also fewer collaborations in general (Fuchs Von Stebut and Allmendinger 2001). Possible explanations for the smaller average professional networks of women and include a greater average amount of family obligations (Long 1990, Wright et al. 2003, Reskin 1978), lower participation of women in international conferences (Lewison 2001, Eisen 2014, Martin 2014), and reduced receipt of research travel funds by women (Bozeman and Corley 2004).

A high, steadily increasing proportion of research papers is written by more than one author (West et al. 2013), making collaboration a key predictor of publication output, and in turn, a key predictor of career prospects (Cohn et al 2014, Jordan et al. 2008, Tower et al. 2007). Additionally, empirical studies suggest that mixed-gender (Campbell et al. 2013) or otherwise diverse (Hong and Page 2004) teams can produce better results on collaborative tasks. For reasons such as these, several studies have tested for gender differences in collaboration frequency or pattern by examining the author lists of published research. To our knowledge, every study of this question has concluded that men publish with other men, and women with women, more often than expected if collaborators assort randomly with regards to gender (Ferber and Teiman, 1980; McDowell and Smith 1992; Ghiasi et al. 2015; Crow and Smykla 2015, Zettler et al. 2016, Fahmy and Young, 2016, Jadidi et al. 2017; Teele and Thelen 2017, Araujo and Fontainha 2017a,b). This preference for same-gendered colleagues is often termed ‘gender homophily’.

However, we believe that all of these studies used sampling strategies that could overstate the strength of homophily, making the true extent of homophily difficult to ascertain from present data. Essentially, whenever samples are taken from two or more discrete sets of literature that vary in the author gender ratio and that are largely not connected by collaboration, the number of same-gendered coauthors will be inflated (\autoref{wahlund_plot}). This can make it appear as though authors preferentially publish with same-gendered colleagues if no gender preferences exist, or even if researchers prefer opposite-gendered colleagues. For example, a sample of bioinformatics and cell biology papers might contain an excess of mostly-male and mostly-female author lists, simply because researchers preferentially work with colleagues from the same discipline, and because more women work in cell biology than in bioinformatics [@Holman_2018]. We name this issue the Wahlund effect, by analogy with the Wahlund effect in population genetics (Wahlund 1928), whereby spatial differences in allele frequencies create an excess of homozygotes across the whole population, even if alleles assort randomly within each subpopulation.

In the present study, we test whether academics tend to co-publish with same-gendered collaborators, while controlling for the Wahlund effect as strictly as possible. Our study uses a recently-published dataset describing the gender of 35.5m authors from 9.15m articles indexed on PubMed [@Holman_2018]. Holman et al. [-@Holman_2018] reported large differences in the gender ratio of authors across research disciplines, journals, countries, and across the years 2002-2016. We therefore tested for gender homophily after restricting the analysis to particular journals (i.e. research specialties), time periods, and countries.

# Methods
## Obtaining author lists and assigning gender
We used the dataset collected from PubMed by Holman et al. [-@Holman_2018]. Briefly, that dataset was created by downloading every single article indexed on PubMed, from the oldest articles up to the present (which was August 2016), and attempting to assign a gender to each author based on their given name, using the web service genderize.io. Each journal was assigned to one of 107 scientific disciplines (e.g. “Nephrology” or “Cell Biology”), using PubMed’s journal categorisations where available. Because the present study focuses on co-authorship, all single-author papers were removed from the dataset. To simplify the statistical analysis, we also discarded all papers for which we could not determine the gender of every author with at least 95% certainty. To mitigate Wahlund effects caused by variation in the gender ratio of researchers over time (see below), we also discarded all papers except those that were published 0-1 or 10-11 years before the PubMed data were collected (i.e. 20 August 2016). Lastly, we excluded journals with fewer than 50 suitable papers. This left us with a dataset from August 2015 - August 2016 containing `r sample.size$n2015[3]` papers with `r sample.size$n2015[4]` authorships, and a dataset from August 2005 - August 2006 containing `r sample.size$n2005[3]` papers and `r sample.size$n2005[4]` authorships. These papers came from `r sample.size$n2015[2]` and `r sample.size$n2005[2]` journals respectively, which were grouped into `r sample.size$n2015[1]` and `r sample.size$n2005[1]` research disciplines. There was a median of `r sample.size$n2015[5]` (`r sample.size$n2005[5]`) papers per journal, `r sample.size$n2015[6]` (`r sample.size$n2005[6]`) authors per journal, and `r sample.size$n2015[7]` (`r sample.size$n2005[7]`) authors per paper (the first number is for 2015-6, and the bracketed number for 2005-6).

## Calculating $\alpha$, the coefficient of homophily
Inspired by Bergstrom et al. (http://www.eigenfactor.org/gender/assortativity/measuring_homophily.pdf), we defined the coefficient of homophily as $\alpha = p - q$, where $p$ is the probability that a randomly-chosen co-author of a _male_ author is a man and $q$ is the probability that a randomly-chosen co-author of a _female_ author is a man. Therefore, $\alpha$ > 0 suggests that same-gender authors publish together more often than expected if authors assort randomly with respect to gender (homophily), $\alpha$ < 0 suggests that opposite-gender authors publish together more often than expected (heterophily), and $\alpha$ = 0 suggests random assortment with respect to gender.

To estimate $\alpha$ for a particular subset of the scientific literature, we estimated $p$ as the average proportion of men’s co-authors who are men (averaged across all papers with at least one man author), and $q$ as the average proportion of women’s co-authors who are men (averaged across all papers with at least one woman author). To estimate the 95% confidence intervals on $\alpha$ for a given set of $n$ papers, we sampled $n$ papers with replacement 1000 times, estimated $\alpha$ on each sample, and recorded the 95% quantiles of the 1000 resulting estimates.

As well as calculating $\alpha$ for all authors, we calculated $\alpha$ for first and last authors only. $\alpha$ was again defined as $p - q$, but this time $p$ was estimated as the average proportion of male co-authors on papers with a male first/last author, and $q$ was estimated as the average proportion of male co-authors on papers with female first/last authors. We did not calculate $\alpha$ for other authorship positions (e.g. second or third authors) because this would necessitate reducing the dataset to papers with a sufficiently long authorship list, complicating comparison with other authorship positions that were calculated using a different sample.

Our test assumes that the expected value of $\alpha$ is zero if authors randomly assort, but for small datasets this assumption is not always true (as pointed out by Carl T. Bergstrom in a blog post, http://www.eigenfactor.org/gender/assortativity/note_to_eisen.rtf). To borrow Prof. Bergstrom’s example, consider a small research specialty comprising just two men and two women researchers, who have together produced six two-author papers: one paper in each of the six possible two-author combinations. For these six papers, $\alpha$ = $-\frac{1}{3}$, even though same- and opposite-gendered coauthors were selected in equal proportion to their frequency in the pool of possible collaborators. To control for the fact that the null expectation for $\alpha$ is not necessarily zero (particularly for very small datasets), we devised an adjusted version of the coefficient of homophily, which we term $\alpha'$. Every time we calculated $\alpha$ for a set of papers, we also determined the expected value of $\alpha$ under the null hypothesis that authors assort randomly with respect to gender. This was accomplished by randomly shuffling authors across papers 1000 times, recalculating $\alpha$, and taking the median. We then calculated $\alpha'$ by subtracting the null expectation for $\alpha$ from the observed value. We also used the null-simulated $\alpha$ values to calculate a two-tailed p-value for the observed value of $\alpha$; the p-value was defined as the proportion of null simulations for which $|\alpha_{null}| > |\alpha_{obs}|$. We applied false discovery rate correction to each set of p-values to account for multiple testing (Benjamini and Hochberg 1995).

As expected, $\alpha'$ was usually almost identical to $\alpha$ (Figure S1), but $\alpha$ was downwardly biased relative to $\alpha'$ for small datasets (Figure S2). Additionally, the correlation between $\alpha'$ and sample size was negligible ($R^2$ < 0.01), suggesting that $\alpha'$ effectively removed the dependence of $\alpha$ on sample size. We therefore used the adjusted statistic $\alpha'$ in all our analyses.

## Minimising the Wahlund effect: discipline and time period
As discussed in the Introduction, the Wahlund effect can give the appearance of gender homophily (reflected in our study as $\alpha'$ > 0) whenever one lumps together different ‘subpopulations’ of academic papers that differ in their author gender ratios (Figure 1).

To minimise bias in $\alpha'$ due to the Wahlund effect, we restricted each set of papers to a single research specialty to the greatest extent allowed by our data. Specifically, we only calculated $\alpha'$ for individual journals, since papers from the same journal typically focus on one or a few closely related topics. Although some journals, e.g. _PLoS ONE_, publish research from diverse disciplines with very different author gender ratios [@Holman_2018], calculating $\alpha'$ for these highly multidisciplinary journals is still useful as a contrast. The difference in $\alpha'$ between highly multidisciplinary and more specialised journals, e.g. _PLoS ONE_ versus _PLoS Computational Biology_, gives a rough estimate of the extent to which multidisciplinarity can inflate $\alpha'$.

As well as varying between disciplines, the gender ratio of authors has changed markedly over time [@Holman_2018]. Because the gender ratio was more male-biased in the past, $\alpha'$ would be inflated if we calculated it for a sample of papers published over a long time frame. To minimise this effect, we only sampled papers from two one-year periods (namely 2005-6 and 2015-16). The median change per year in % (fe)male authors across journals is < 0.5% [@Holman_2018], and so restricting our dataset to a single year should prevent temporal changes in gender ratio from noticeably inflating our estimates of $\alpha'$.

## Minimising the Wahlund effect: author country of affiliation
A Wahlund effect could arise even if one calculates $\alpha'$ for a single discipline and time period, because of variation in the gender ratio of researchers from different countries. For example, Holman et al. [-@Holman_2018] showed that authors based in Serbia are more than twice as likely to be women as are authors based in Japan. Therefore, a dataset containing a mix of papers from teams of authors based in these two countries would probably contain an excess of same-sex coauthorships, even if collaboration were random with respect to gender within each country.

To address this, we also analysed every combination of journal and author country of affiliation for which we had enough data (i.e. 50 or more papers published in 2015-16). For simplicity, we restricted the dataset to only include papers for which Holman et al. [-@Holman_2018] had successfully identified the country of affiliation for all authors on the paper, and all authors shared the same country of affiliation. Restricting the dataset in this fashion produced enough data to measure $\alpha'$ for 325 combinations of journal and country (median: 70 papers and 273 authors per combination).

## Calculating standardised journal impact factor
We obtained the 3-year impact factor for each journal from Clarivate Analytics. To account for large differences in impact factor between disciplines, we took the the residuals from a model with $Log_{10}$ impact factor as the response and the research discipline of the journal as a random effect. Thus, journals with a positive standardised impact factor have a higher mean number of citations than the average for journals in their discipline. We then used Spearman rank correlation to test whether our estimates of gender homophily covaried with journal impact factor.

## Statistical analysis
Previous authors (e.g. Bonham & Stefan 2017) have hypothesised that senior scientists preferentially recruit staff and students of the same gender, and/or that junior researchers preferentially select same-gendered mentors. In the majority of disciplines, authorship conventions mean that the first-listed author is often an early-career researcher, while the author listed last is more likely to be a senior researcher leading a research team (Wren et al. 2007). Assuming that senior researchers are the main drivers of homophily and that there are enough papers with three or more authors, we predict that the last author’s gender will be the strongest predictor of the remaining authors’ genders (i.e. the gender of the last author will be more salient than that of the first author, or any other authorship position). This is because the first author’s gender would simply be an imperfect correlate of the true causal effect, while the last author’s gender would be the causal effect itself.

To test whether $\alpha'$ for last authors tends to be higher than $\alpha'$ for first authors for any given dataset, we used a linear mixed model implemented in the ``lme4`` and ``lmerTest`` packages for R, with _authorship position_ (first or last) as a fixed factor, and _journal_ and _research discipline_ as crossed random effects. The response variable was $\alpha'$, and we weighted each observation by the inverse of the standard error from our estimate of $\alpha'$, meaning that more accurate measurements of $\alpha'$ had more influence on the results. We used a similar model to test for a difference in $\alpha'$ between the 2005-6 and the 2015-16 datasets, with two differences: instead of authorship position, we fit year range as a two-level fixed factor, and used $\alpha'$ estimated for all authors (not first/last authors) as the response variable.

The relationship between the gender ratio of authors publishing in a journal and its $\alpha'$ value appeared nonlinear (Figure 4). We therefore fit a generalised additive model with thin plate regression spline smoothing, implemented using the ``mgcv`` package for R.

# Results
## Gender homophily by discipline, time period, and authorship position
\autoref{alpha_histograms} shows the distribution of $\alpha'$ estimates in 2015-2016 across all journals for which we recovered sufficient data, when $\alpha'$ was calculated for all authors, first authors only, or last authors only. The great majority of journals had $\alpha'$ > 0, and for many of these the corrected two-tailed p-values suggested that $\alpha'$ was significantly greater than zero (1469/2077 journals were significant in 2015-16, and 404/1192 in 2005-6). Only 2/2077 journals had statistically significantly heterophily (i.e. $\alpha'$ < 0) in 2015-16, and 1/1192 in 2005-6. The remaining 606 or 787 journals (in 2015 and 2005 respectively) had a value of $\alpha'$ not significantly different from zero, such that we could not reject the null hypothesis of random assortment with respect to gender.

$\alpha'$ was significantly higher in the sample from 2015-16 relative to 2005-6, but the difference in means was small (Figure S3; Effect of the fixed factor ‘time period’ in a linear mixed model of the data for all author positions: Cohen’s $d$ = 0.09 ± 0.04, $t_{948}$ = 2.51, p = 0.012). This result illustrates that gender homophily was slightly stronger in 2015-6 relative to 2005-6.

When comparing pairs of $\alpha'$ values estimated for the first and last authors for the same journals, we found that $\alpha'$ tended to be higher for first authors than for last authors (Figure S4; Effect of the fixed factor ‘Authorship position’ in a linear mixed model: Cohen’s d = 0.07 ± 0.02, $t_{1988}$ = 4.48, p < 0.0001). This suggests that the gender of the first author was a slightly stronger predictor of the remaining authors’ genders than the gender of the last author, i.e. the opposite of our prediction based on the hypothesis that senior scientists are causally responsible for homophily.

## Possible differences in the strength of homophily between disciplines
\label{alpha_gender_ratio} illustrates the variance in journal homophily values ($\alpha'$) across scientific disciplines. All disciplines had a positive average $\alpha'$, although homophily appeared somewhat stronger in some disciplines than others (e.g. mean $\alpha'$ was 0.12 ± 0.02 for Urology journals, and 0.03 ± 0.01 for Veterinary Medicine journals; Table S3). However, there was little evidence for strong differences in $\alpha'$ between disciplines: the random factor ‘discipline’ explained at most 1% of the variance in $\alpha'$ in the two linear mixed models described in the previous section (see mixed models in Online Supplementary Material). This implies that the processes causing positive $\alpha'$ are similarly strong in all the disciplines we examined.

Reassuringly, we found no support for our prediction that journals publishing on a wide range of topics should have high $\alpha'$ values due to the Wahlund effect \label{alpha_gender_ratio}. For example, the journal category “Multidisciplinary” - which includes journals like PLoS ONE, Nature, Science, and PNAS - does not have notably elevated $\alpha'$. This result suggests that our estimates of homophily were not greatly inflated by the presence of disparate research topics (with variable author gender ratios) in the same journal.

## Relationship between gender homophily and gender ratio
One might predict that researchers would be more likely to seek out same-gendered colleagues in strongly gender-biased disciplines (e.g. Surgery or Nursing), relative to disciplines with a comparatively gender-balanced workforce (e.g. Psychiatry). However, we instead found evidence of the reverse: there was a positive, non-linear relationship between the overall gender ratio of all authors publishing in a particular journal [as estimated by @Holman_2018], and the estimated value of $\alpha'$ for all authors and for first authors \label{alpha_country}. That is, journals with a balanced or female-biased author gender ratio tended to have higher $\alpha'$ than journals with a male-biased author gender ratio (GAM smooth terms p < 0.001, see Online Supplementary Material). The relationship was not statistically significant when $\alpha'$ was calculated for last authors (GAM, p = 0.142), though the trend appeared similar \label{alpha_country}.

## Relationship between journal impact factor and gender homophily
We observed a noisy but statistically significant linear relationship between discipline-standardised journal impact factor and $\alpha'$, such that journals with a high impact factor for their discipline had weaker gender homophily than did journals with a low impact factor for their discipline (Figure 5; linear regression: $R^2$ = `r R2`, $t_{1415}$ = -8.0, p < 0.0001).

## Analysis correcting for differences between countries
When we restricted the analysis to only include authors with affiliations from a single country, we found statistically significant homophily for 72 of the 325 journal-country combinations tested (64 unique journals and 18 unique countries), and no significant evidence of heterophily (Figures S5-S6). Additionally, the values of $\alpha'$ calculated for each journal-country combination were mostly very similar to the $\alpha'$ values calculated for the journal as a whole (i.e. when pooling papers from different countries); the average difference in $\alpha'$ was 0.002 (Figure S7). These results suggest that our main analysis’ findings of widespread homophily are not driven by a Wahlund effect resulting from gender differences between countries.

# Discussion
We found evidence that researchers tend to publish with same-gendered coauthors, even after implementing stringent controls for ‘spurious’ homophily resulting from Wahlund effects (\autoref{wahlund_plot}). Our data therefore reaffirm earlier conclusions (Ferber and Teiman 1980; McDowell and Smith 1992; Crow and Smykla 2015; Ghiasi et al. 2015; Zettler et al. 2016; Fahmy and Young 2016; Jadihi et al. 2017; Araujo and Fontainha 2017a; Teele and Thelen 2017) that research collaborations occur between same-gendered colleagues substantially more often than expected if collaborators are selected without regard to their gender. Many journals had $\alpha'$ values over 0.1, meaning that the gender ratio of the average man's coauthors is skewed towards men by >10%, relative to the gender ratio of a woman's coauthors (or equivalently, that the gender ratio of women's coauthors is >10% more female than that of men's). Few journals had $\alpha'$ values below zero, and fewer still had significantly negative $\alpha'$ values, suggesting that researchers rarely seek out collaborators of the opposite gender (or at least, that they do so less often than they seek out same-gendered colleagues).

We found no evidence that gender homophily is restricted to particular disciplines: instead, there was evidence of homophily across many diverse research disciplines. We also found that gender homophily was marginally stronger in 2015-2016, relative to 2005-2006. This result suggests that recent increases in the number of women working in STEMM, as well as recent efforts to recruit, retain, and celebrate women in STEMM, have (if anything) strengthened gender homophily rather than weakening it.

Gender homophily was slightly stronger in research disciplines with a female-biased gender ratio, as compared to male-biased or gender-balanced fields. One interpretation of this result is that men are more likely to seek out other men in fields in which men are in the minority, relative to the homophily shown by women in fields in which women are in the minority.

We also found that journals with a high impact factor (relative to other journals in the same research discipline) tended to have weaker gender homophily than did low-impact journals. If one accepts the contentious claim that papers published in high-impact journals tend to be of higher quality than those in low-impact journals [e.g. @Garfield_2006], then this result provides correlational support for the hypothesis that mixed-gender teams produce better research than single-gender teams (Campbell et al. 2013).

Contrary to our expectations, we found no evidence that homophily is driven mostly by senior academics, and no evidence that homophily is strongest in fields where women are in the minority. Interestingly, homophily was negatively correlated with journal impact factor (standardised by research discipline), as predicted if mixed-gender teams produce better research.

Academic publishing is one of the most (if not the most) important tool for a researcher to display their work, and is frequently used as a success indicator for being promoted, hired as a faculty member, or being invited as plenary speaker. Our finding that researchers are more likely to publish with same-gendered coauthors implies that co-authorship listing is not a random process, but instead creates additional gender imbalance. Homophily could make it harder for women to find collaborators in disciplines where women are in the minority [i.e. almost all STEMM fields; @Holman_2018]. Because women make up for as little as 30% of the STEMM scientist community (REF), they are therefore less likely to be picked as coauthors, and, as a result, gender homophily creates further gender imbalance and disadvantages for women. Men tend to publish more in high impact journals (REF) as first author, and generate more citations than papers lead by first author women (REF). As a consequence, gender homophily is also likely to contribute to lower publication rates for women (McDowell and Smith 1992), and lower scientific success (REF).

Why does gender homophily exists in science? Collaboration often leads to co-authorship publications, citations and additional professional recognitions, therefore first author should always be considering collaborators that are most likely to enhance their productivity and in the long term also their academic success, rather than pick co-authors according to their gender. First of all, it is likely that decades ago, the probability to meet a woman colleague working on the same area of expertise was slimmer than now. However, the number of women in STEM is increasing, and it is not likely to be an obstacle to collaboration nowadays.

# Acknowledgements
We are grateful to XXXX.

# References
<div id="refs"></div>

\newpage
# Figures
![The Wahlund effect can make it appear as if authors prefer to publish with same-gendered colleagues, even if no such preference exists. Here, coloured circles represent male and female authors, and coauthors are linked with lines. Across the whole set of ten papers, there is an apparent excess of same-gender collaborations. Specifically, there are six same-gender papers and only four mixed-gender papers, which is fewer than the $10\times2\times0.5\times0.5 = 5$ mixed-gender papers we would expect under the null hypothesis that authors assort randomly with respect to gender. However, within each subset, there is no evidence that authors prefer to publish with same-gendered individuals. The Wahlund effect will tend to inflate the frequency of same-sex coauthors whenever the data is composed of two or more disconnected subsets of literature with different author gender ratios; these subsets could be research disciplines, older versus newer papers, or papers from authors in different countries. \label{wahlund_plot}](../figures/Figure 1 - Wahlund figure.pdf)

\newpage
![Of the `r n.journals` journals for which we had adequate data in 2015-2016, `r n.sig.homo` showed statistically significant evidence of homophily (denoted by $\alpha' > 0$), and `r n.sig.hetero` showed statistically significant evidence of heterophily ($\alpha' < 0$), after adjusting p-values using Benjamini-Hochberg false discovery rate correction. The white area shows the number of journals for which homophily was significantly stronger than expected under the null hypothesis (p < 0.05), while the blue area shows all the remainder. Patterns were similar whether $\alpha'$ was calculated for all authors, for first authors only, or for last authors only. \label{alpha_histograms}](../figures/figure2.pdf)

\newpage
![There is a weakly positive, non-linear relationship between the gender ratio of authors publishing in a journal, and the coefficient of homophily ($\alpha'$). Specifically, journals with 50% women authors or higher tended to have more same-sex coauthorships than did journals with predominantly men authors.  This relationship held whether $\alpha'$ was calculated for all authors, first authors only, or last authors only. A negative value on the x-axis denotes an excess of men authors, a positive value denotes an excess of women authors, and zero denotes gender parity. The lines were fitted using generalised additive models with the smoothing parameter $k$ set to 3. \label{alpha_gender_ratio}](../figures/figure3.pdf)

\newpage
![Journal impact factor (expressed relative to the average for the discipline) is negatively correlated with $\alpha'$. The relationship is noisy ($R^2$ = `r R2`), but the results suggest that journals with strong homophily tend to have lower impact factors than journals with weak homophily in the same discipline. \label{impact_factor}](../figures/figure5.pdf)