### 2019-04-11 manuscript/homophily_manuscript.tex

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Researchers collaborate with same-gendered colleagues more often than expected across the life sciences},
            pdfauthor={Luke Holman* and Claire Morandin; *;},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Researchers collaborate with same-gendered colleagues more often than
expected across the life sciences}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Luke Holman* and Claire Morandin\textsuperscript{§} \\ *\textit{luke.holman@unimelb.edu.au} \\ \textsuperscript{§}\textit{claire.morandin@helsinki.fi} \vspace{5mm}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath} \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}

\pretitle{\begin{flushleft}\LARGE}
\posttitle{\par\end{flushleft}\vskip 0.5em}
\preauthor{\begin{flushleft}\large \lineskip 0.5em}
\postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
\postdate{\par\end{flushleft}}

\usepackage{fancyhdr}

\pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
\rhead[]{Gender and coauthorship}

\usepackage[noabbrev,capitalise]{cleveref} \usepackage{titlefoot} \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno}

\linenumbers

\begin{document}
\maketitle
\begin{abstract}
Evidence suggests that women in academia are hindered by conscious and
unconscious biases, and often feel excluded from formal and informal
opportunities for research collaboration. In addition to ensuring
fairness and helping to redress gender imbalance in the academic
workforce, increasing women's access to collaboration could help
scientific progress by drawing on more of the available human capital.
Here, we test whether researchers tend to collaborate with same-gendered
colleagues, using more stringent methods and a larger dataset than in
past work. Our results reaffirm that researchers co-publish with
colleagues of the same gender more often than expected by chance, and
show that this `gender homophily' is slightly stronger today than it was
10 years ago. Contrary to our expectations, we found no evidence that
homophily is driven mostly by senior academics, and no evidence that
homophily is stronger in fields where women are in the minority.
Interestingly, journals with a high impact factor for their discipline
tended to have comparatively low homophily, as predicted if mixed-gender
teams produce better research. We discuss some potential causes of
gender homophily in science. \vspace{5mm}

\par

\noindent \textbf{Keywords:} Gender bias, Homophily, Research
collaboration, Text mining, Women in STEM.
\end{abstract}

\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
\maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Women are severely underrepresented in many branches of science,
technology, engineering, mathematics, and medicine (STEMM), and face
additional challenges and inequities relative to men {[}1--5{]}. On
average, women occupy more junior positions {[}6,7{]} with lower
salaries {[}8,9{]}, receive less grant money {[}10,11{]}, are promoted
more slowly {[}12--15{]}, and are allocated fewer resources {[}16{]} and
less research funding {[}17--19{]}. Experimental evidence suggests that
bias against women plays a major role in generating these differences
{[}20,21{]}.

Writing papers, networking, and collaboration are all instrumental to
research productivity and academic career advancement {[}22--25{]}, and
dozens of studies have tested for gender differences in these areas
{[}5,26--29{]}. For example, studies have concluded that women tend to
be less involved in international collaboration {[}19,28,30--32{]},
collaborate less within their own university departments {[}31{]}, have
less prestigious collaborations {[}33{]}, and fewer collaborations in
total {[}34{]}. These gender differences in collaboration presumably
have multiple causes, which might include implicit and explicit gender
bias {[}20{]}, differential family obligations {[}33,35,36{]}, gender
differences in confidence or self-esteem {[}37{]}, concerns relating to
sexual harassment {[}38{]}, and unequal access to conferences {[}39{]}
or travel funds {[}32{]}.

A high, steadily increasing proportion of research papers is written by
more than one author {[}3{]}, making collaboration a key predictor of
publication output, and thus of career prospects {[}40,41{]}.
Additionally, empirical studies imply that mixed-gender or otherwise
diverse teams produce better outputs on collaborative tasks than less
diverse teams {[}42--48{]}. For reasons such as these, multiple studies
have examined the author lists of published research articles in order
to test for gender differences in collaboration frequency or pattern. To
our knowledge, most or all such studies imply that men co-publish with
men, and women with women, more often than expected if collaborators
assort randomly with respect to gender {[}49--58{]}. This non-random
assortment is often termed `gender homophily'.

We believe that most or all earlier studies of gender homophily were
hindered by a largely unacknowledged statistical issue that we will
refer to as the Wahlund effect (Fig 1), by analogy with the conceptually
similar Wahlund effect in population genetics {[}59{]}. The Wahlund
effect makes it deceptively difficult to test for gender-based co-author
choice simply by counting the relative number of same- and mixed-gender
coauthorships. Essentially, the Wahlund effect means that whenever
coauthorship data are sampled from two or more discrete sets of
literature, which vary in the author gender ratio and which are largely
unconnected by collaboration, the number of same-gendered coauthors will
be inflated. This can give the impression that authors preferentially
publish with same-gendered colleagues even if no gender preferences
exist, or if the true preference is for opposite-gendered colleagues
(`gender heterophily'). For example, a sample of literature containing a
mixture of bioinformatics and cell biology papers will probably contain
an excess of mostly-male and mostly-female author lists, simply because
researchers usually collaborate within their own discipline, and because
the author gender ratio is more male-biased in bioinformatics than in
cell biology {[}5{]}.

\vspace{2cm}

\textbf{Fig 1}: The Wahlund effect can make it appear as if authors
publish with same-gendered colleagues disproportionately often, even if
collaboration is completely random with respect to gender. Here,
coloured circles represent male and female authors, and coauthors are
linked with lines. Across the whole set of ten papers, there is an
apparent excess of same-gender collaborations: there are six same-gender
papers and only four mixed-gender papers, which is fewer than the
\(10\times2\times0.5\times0.5 = 5\) mixed-gender papers expected under
the null hypothesis that authors assort randomly. However, within each
subset, there is no evidence that authors prefer to publish with
same-gendered individuals (if anything, this small dataset suggests
gender heterophily). The Wahlund effect will tend to inflate the
frequency of same-gender coauthorships whenever the data is composed of
two or more disconnected subsets of literature with different author
gender ratios; these subsets could be research disciplines, older versus
newer papers, or papers from authors based in different countries. The
example countries and disciplines were selected based on data in
{[}5{]}. \vspace{2cm}

In the present study, we test whether life sciences researchers tend to
co-publish with same-gendered colleagues, while controlling for the
Wahlund effect as strictly as we are able. We use a recently-published
dataset describing the gender of 35.5 million authors from 9.15 million
articles indexed on PubMed {[}5{]}. Holman et al. {[}5{]} reported large
differences in the gender ratio of authors across research disciplines,
journals, countries, and across the years 2002-2016. We therefore tested
for gender homophily while restricting our analysis to particular
journals (a proxy for research specialties), time periods, and
countries. We quantified gender assortment using a metric called
\(\alpha'\) {[}60{]}, which is positive when same-gender authors publish
together more often than expected (gender homophily), negative when
opposite-gender authors publish together more often than expected
(heterophily), and equal to zero when authors assort randomly with
respect to gender (see Methods).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{gender-homophily-by-discipline-time-period-and-authorship-position}{%
\subsection{Gender homophily by discipline, time period, and authorship
position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}}

Fig 2 shows the distribution of \(\alpha'\) estimates in 2015-2016
across all journals for which we recovered sufficient data, when
\(\alpha'\) was calculated for all authors, first authors only, or last
authors only. Most journals had positive values of \(\alpha'\) (77-92\%,
depending on time period and author type; S1 Data), and for many of
these the false discovery rate (FDR)-corrected p-values suggested that
\(\alpha'\) was significantly greater than zero (1469/2077 journals were
significant in 2015-16, and 404/1192 in 2005-6; S1 Data). Only 2/2077
journals had statistically significant heterophily (i.e.
\(\alpha' < 0\)) in 2015-16, and 1/1192 in 2005-6 (S2 Table). The
remaining 606 or 787 journals (in 2015 and 2005 respectively) had a
value of \(\alpha'\) not significantly different from zero, consistent
with the null hypothesis of random assortment with respect to gender. We
also confirmed that in most journals (S2 Data) and most research
disciplines (S3 Data, S1 Fig), the majority of papers had multiple
authors.

\vspace{2cm}

\textbf{Fig 2}: Of the 2116 journals for which we had adequate data in
2015-2016, 825 showed statistically significant evidence of gender
homophily (denoted by \(\alpha' > 0\)), and 1 showed statistically
significant evidence of heterophily (\(\alpha' < 0\)), after false
discovery rate correction. In the stacked density plot, the white area
shows the number of journals for which homophily was significantly
stronger than expected under the null hypothesis (corrected p
\textless{} 0.05), while the blue area shows all the remainder. Patterns
were similar whether \(\alpha'\) was calculated for all authors, for
first authors only, or for last authors only. Points in the right panel
show \(\alpha'\) for individual journals. \vspace{2cm}

\(\alpha'\) was significantly higher in the literature sample from
2015-16 relative to 2005-6, though the difference in means was small (S2
Fig; Effect of the fixed factor `Time period' in a linear mixed model of
the data for all author positions: Cohen's \(d\) = \(0.091{\pm}0.04\),
\(t_{953}\) = 2.42, p = 0.016).

When comparing pairs of \(\alpha'\) values estimated for the first and
last authors for the same journals, we found that \(\alpha'\) tended to
be higher for first authors than for last authors (S3 Fig; Effect of the
fixed factor `Authorship position' in a linear mixed model: Cohen's
\(d\) = \(0.065{\pm}0.02\), \(t_{2024}\) = 4.28, p \textless{} 0.0001).
This suggests that the gender of the first author was a slightly
stronger predictor of the remaining authors' genders than the gender of
the last author, i.e.~the opposite of what is predicted if senior
scientists are causally responsible for homophily.

\hypertarget{variance-in-homophily-between-disciplines}{%
\subsection{Variance in homophily between
disciplines}\label{variance-in-homophily-between-disciplines}}

Fig 2 illustrates the variance in journal homophily values (\(\alpha'\))
across scientific disciplines. All disciplines had positive mean
\(\alpha'\) (averaged over journals), although homophily appeared
somewhat stronger in some disciplines than others (e.g.~mean \(\alpha'\)
was \(0.12{\pm}0.02\) for Urology journals and \(0.03{\pm}0.01\) for
Veterinary Medicine journals; Fig 2, S4 Data). However, there was no
formal evidence for consistent differences in \(\alpha'\) between
disciplines: the random factor `Discipline' explained around 1\% of the
variance in \(\alpha'\) in the two linear mixed models described in the
previous section (see Fig 2 and mixed models in Online Supplementary
Material). Thus, the causal mechanisms underlying the observed positive
\(\alpha'\) values appear to be similarly strong in all the disciplines
we examined.

There was no indication that journals publishing on a wide range of
topics have higher \(\alpha'\) values than more specialised journals due
to the Wahlund effect (Fig 1). For example, the journal category
`Multidisciplinary' -- which includes general interest journals like
\emph{PLOS ONE}, \emph{Nature}, \emph{Science}, and \emph{PNAS} -- did
not have markedly elevated \(\alpha'\) (Fig 2). This result suggests
that our estimates of homophily, and estimates from some of the earlier
studies of homophily listed in the Introduction, are probably not
markedly inflated by the presence of disparate research topics (with
variable author gender ratios) being published within individual
journals.

Nevertheless, when we calculated \(\alpha\) across all non-single-author
papers in our entire 15-year PubMed dataset (as before, excluding papers
where at least one author's gender was unknown; n = \textgreater{}3
million papers, \textgreater{}16 million authors), we found that
\(\alpha\) was 0.126. This figure is almost double the median value of
\(\alpha'\) for individual journals (Fig 2; \(\alpha' = 0.070\) for `All
authors'), suggesting that lumping together papers from different fields
and different time periods can indeed produce spurious evidence for
gender homophily as outlined in Fig 1.

\hypertarget{relationship-between-gender-homophily-and-number-of-authors}{%
\subsection{Relationship between gender homophily and number of
authors}\label{relationship-between-gender-homophily-and-number-of-authors}}

Papers with two authors had significantly lower (but still positive)
\(\alpha'\) values relative to papers with more than two authors, while
papers with 3, 4 or \(\ge5\) authors had essentially identical average
\(\alpha'\) values (Fig 3). Specifically, the posterior estimate of mean
\(\alpha'\) was 0.014 (95\% CIs: 0.002 - 0.026) for 2-author papers and
0.065 (95\% CIs: 0.056 - 0.074) for 3-author papers (and roughly the
same for 4- and \(\ge5\)-author papers; Fig 3). One possible explanation
for this finding is that 2-authors papers are more likely to have an
author list that is evenly split between career stages (e.g.~a
postgraduate student and their supervisor), increasing the chance that
the authors are mixed gender (see section `Theoretical expectations for
\(\alpha\) when the gender ratio differs between career stages'). The
result also suggests that the causal mechanisms responsible for gender
homophily are similar in small (e.g.~3-author) and larger (\(\ge5\)
author) collaborations (and across disciplines where small versus large
collaborations are the norm).

\vspace{2cm}

\textbf{Fig 3}: The coefficient of homophily (\(\alpha'\)) was slightly
less positive when calculated for two-author papers only, relative to
papers with longer author lists. The individual points, whose
distribution is summarised by the violin plots, correspond to individual
journals. The larger white points show the mean for each group (and its
95\% CIs), as calculated by a Bayesian meta-regression model accounting
for repeated measures of \(\alpha'\) within journals, as well as the
precision with which \(\alpha'\) was estimated. \vspace{2cm}

\hypertarget{relationship-between-gender-homophily-and-gender-ratio}{%
\subsection{Relationship between gender homophily and gender
ratio}\label{relationship-between-gender-homophily-and-gender-ratio}}

We next tested whether researchers are more or less likely to publish
with same-gendered colleagues in strongly gender-biased disciplines
(e.g.~Surgery or Nursing), relative to disciplines with a comparatively
gender-balanced workforce (e.g.~Psychiatry). We found a positive,
non-linear relationship between the gender ratio across all the authors
publishing in a particular journal {[}5{]}, and the estimated value of
\(\alpha'\) for all authors and for first authors, but not last authors
(S4 Fig). Journals with a balanced or female-biased author gender ratio
tended to have higher \(\alpha'\) (i.e.~stronger homophily) than
journals with a male-biased author gender ratio (GAM smooth term p =
0.0002 for all-author homophily, p \textless{} 0.0001 for first-author
homophily, and p = 0.13 for last-author homophily).

\hypertarget{relationship-between-journal-impact-factor-and-gender-homophily}{%
\subsection{Relationship between journal impact factor and gender
homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}}

We observed a noisy but statistically significant linear relationship
between standardised journal impact factor and \(\alpha'\), such that
journals with a high impact factor for their discipline had weaker
gender homophily than did journals with a low impact factor for their
discipline (Fig 4; linear regression: \(R^2\) = 0.043, \(t_{1415}\) =
-8.0, p \textless{} 0.0001). The slope of the regression was
\(-0.012{\pm}0.0015\), indicating that increasing the
discipline-standardised impact factor by one standard deviation is
associated with a reduction in \(\alpha'\) of 0.012. The Spearman
correlation coefficient was -0.19 (p \textless{} 0.0001).

\vspace{2cm}

\textbf{Fig 4}: Journal impact factor (expressed relative to the average
for the discipline) is negatively correlated with \(\alpha'\). The
relationship is noisy (\(R^2\) = 0.043, Spearman correlation = -0.19),
but the results indicate that journals with strong gender homophily tend
to have lower impact factors than journals from the same discipline that
have weaker homophily. \vspace{2cm}

\hypertarget{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}{%
\subsection{Analysis accounting for differences in author gender ratio
between
countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}}

When we restricted the analysis by country, we observed statistically
significant homophily for 72 of the 325 journal-country combinations
tested (64 unique journals and 18 unique countries), and no significant
heterophily (S5-S6 Fig). Additionally, the values of \(\alpha'\)
calculated for each journal-country combination were only very slightly
lower than the \(\alpha'\) values calculated for the journal as a whole
(i.e.~when pooling papers from different countries, as was done to make
Fig 2): on average, the difference in \(\alpha'\) was only 0.002 (S7
Fig). These results suggest that our findings of widespread homophily in
the main analysis were not driven solely by a Wahlund effect resulting
from gender differences between countries.

\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}{%
\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}}

Given that we cannot identify individual researchers or their career
stages, we used a simple model to derive the theoretical expectations
for \(\alpha\) when the gender ratio differs between career stages (see
Methods). As shown in Fig 5, we predict that \(\alpha\) is expected to
be non-zero, even if collaborators are randomly selected with respect to
gender, provided that there is a gender gap between career stages. The
extent to which \(\alpha\) deviates from zero depends on the relative
frequencies of collaboration within and between career stages (rows and
columns in Fig 5), and the size of the gender gap between stages (\(x\)-
and \(y\)-axes in Fig 5). When \textgreater{}50\% of coauthor pairs
comprise one early-career and one established researcher, we expect
gender heterophily (\(\alpha < 0\)) whenever the gender ratio differs
between career stages. Conversely, when \textgreater{}50\% of
collaborations are between people at the same career stage, we expect
gender homophily (\(\alpha > 0\)). In a few parameter spaces (shown in
red; Fig 5), \(\alpha\) was quite high, and overlapped with the values
that we estimated (Fig 2).

\vspace{2cm}

\textbf{Fig 5}: When the gender ratio of early-career researchers is not
equal to the gender ratio among established researchers, the null
expectation for \(\alpha\) is not necessarily zero. Specifically, if
most collaborations occur between career stages, there will be an excess
of mixed-gender collaborations (\(\alpha < 0\), blue areas), while if
most collaborator pairs comprise two people at the same career stage,
there will be an excess of same-gender collaborations (\(\alpha > 0\),
red areas). However, the conditions required for strong gender homophily
(i.e.~the red areas) are quite restrictive, making it unlikely that this
issue can fully explain the homophily observed in our study.
Additionally, in research disciplines where between-career stage
collaboration is common and there is a shortage of women among
established researchers (i.e.~the blue areas), our study will
underestimate the strength of gender homophily. Contour lines mark
increments of 0.1. \vspace{2cm}

Despite this overlap, Fig 5 suggests that our main conclusions (and
those of other studies of gender homophily) are probably robust to this
career stage issue. We only expect strongly positive \(\alpha\) when A)
the gender ratio is highly skewed across career stages (e.g.~a 5-fold
difference), and B) collaborations between early and established
researchers are very rare (e.g. \textless{}10\% of the total). Both of
these conditions seem unlikely to be true for most fields: the gender
gap across careers stages is generally less pronounced {[}1,5{]}, and it
is very common for early-career researchers to co-publish with an
established mentor {[}61{]}. However, one can get \(\alpha > 0\) for
realistic combinations of parameters, e.g.~a moderate shortage of women
in senior positions coupled with a moderate excess of within-career
stage collaboration, suggesting this effect might contribute to some of
the homophily observed by this and previous studies.

Lastly, we note that if there is a gender gap between career stages and
coauthorships between early-career and established researchers comprise
\textgreater{}50\% of the total, then the baseline expectation for
\(\alpha\) is actually less than zero (blue areas in Fig 5). Therefore,
it is possible that researchers preferentially assort with same-gendered
collaborators even more strongly than implied by our results, at least
for certain journals or research disciplines.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

We found evidence that researchers work with same-gendered coauthors
more often than expected under the null model, even after implementing
stringent controls for Wahlund effects of the kind illustrated in Fig 1.
Our study therefore reaffirms earlier studies' conclusions
{[}49--57,62{]} using stricter methodology, and generalises their
results across the life sciences. Relatively few journals had
\(\alpha'\) values below zero, and almost no journals showed
statistically significant gender heterophily after controlling for
multiple testing. The excess of same-gender coauthorships was quite
large: many journals had \(\alpha' > 0.1\), indicating that the gender
ratio of men's and women's coauthors differs by \textgreater{}10\% in
absolute terms. In relative terms, our findings are even more striking:
for example, if men have 20\% female coauthors and women have 30\% (i.e.
\(\alpha' = 0.1\) in a field with a typical gender ratio {[}5{]}), then
women publish with women 50\% more often than men do.

An important limitation of our study is that we cannot reliably
determine the cause(s) of the observed excess of same-gender
coauthorships. As well as the obvious interpretation -- conscious or
unconscious selection of same-gendered collaborators by men, women, or
both genders -- our results could be partly explained by uncontrolled
Wahlund effects. However, we suspect the contribution of these
uncontrolled artefacts to be minor, for four reasons: we found positive
\(\alpha'\) after controlling for three obvious sources of Wahlund
effect; there was no inflation of \(\alpha'\) in highly
multidisciplinary journals relative to specialised journals; restricting
the data by country yielded similar estimates of \(\alpha'\); and our
modelling work suggested that differences in gender ratio between career
stages are unlikely to fully explain our results. On balance, we believe
the data suggest that it is likely that some researchers preferentially
select same-gendered collaborators, although it is difficult to
ascertain what proportion of people show such a preference, or how much
the strength of the preference varies between individual researchers. We
also note that even in a world in which everyone selected their
collaborators at random with respect to gender, a high proportion of
individual researchers would have entirely same-gendered collaborators
by chance alone (especially in gender-biased disciplines); thus,
individuals who only have same-gendered co-authors are not necessarily
doing anything differently from people with gender-balanced co-authors.

We hypothesised that disciplines with a strongly skewed gender ratio
might show the strongest gender homophily, e.g.~because being in the
minority might increase one's motivation to seek out same-gendered
colleagues. Contrary to this hypothesis, we found no evidence that
gender homophily is restricted to particular disciplines: \(\alpha'\)
was similarly high across the board (Fig 2). Interestingly, gender
homophily was weakest for journals with a male-biased author gender
ratio, and strongest in journals with a female-biased author gender
ratio. One possible reason is that men are more likely to preferentially
seek out male collaborators in fields where men are a minority, relative
to the homophily displayed by women in fields where women are a
minority. However, this latter result only has tentative statistical
support since our sample contains few journals in which most authors are
women (S4 Fig).

We also found that gender homophily was marginally stronger in 2015-2016
relative to 2005-2006. Although this trend might reflect a change in the
gender preferences of researchers seeking collaborators, there are
alternative (and perhaps more likely) explanations. For example, this
trend might result from the increasing number of women working in senior
positions in STEMM over the past decade {[}63--65{]}. As shown in Fig 5,
if enough coauthorships are between junior and senior researchers, a
large gender gap between career stages can give the appearance of
heterophily. As this gender gap between career stages lessens, the
observed values of \(\alpha'\) may increase.

Regarding our finding of weaker homophily among 2-author papers, we
suspect that many 2-author teams comprise a student/postdoc and a senior
staff member, making these teams especially likely to be mixed-gender,
due to the greater shortage of women among senior researchers {[}1,5{]}.
Assuming this interpretation is correct, this result suggests that our
reported \(\alpha'\) values may underestimate the strength of peoples'
preferences for same-gendered collaborators; essentially, women seeking
a senior collaborator could be constrained to work mostly with men,
meaning that people's ideal and realised gender preferences would be
mismatched. On a related note, Ghiasi et al. {[}51{]} argue that women
in engineering are ``compliant {[}in reproducing{]} male-dominated
scientific structures'' because they do not collaborate often enough
with other women (for reference, Figure 7 in {[}51{]} implies that
coauthorships involving two women are \(c\). 30\% more frequent than
expected under random assortment). By contrast, we feel that it may be
counter-productive to recommend that women collaborate primarily with
other women, e.g.~because this constrains women's options (particularly
in fields like the one studied by Ghiasi et al. -- engineering -- where
90\% of professors are men {[}1{]}). Instead, we suggest that
researchers of both genders can help to close the gender gap in STEMM.
In the context of collaboration, one way to do this is to undertake
self-examination to ensure that one is not inadvertently overlooking or
excluding women among potential students and colleagues. One should also
take care to treat male and female collaborators equally, e.g.~in terms
of training and mentoring, allocation of work, and how one descibes the
collaboration to other people (e.g.~in conference presentations, on the
lab website, or in the `Author contributions' section of a paper).
Experimental work suggests that unconscious bias causes people to
undervalue women's research achievements {[}20{]}, and a study of author
contribution statements found observational evidence that menial or
under-valued tasks are more often assigned to women while more
prestigious tasks are assigned to men {[}61{]}.

Our study begs two questions: what causes gender homophily in science,
and are our results cause for concern? We believe that the answers to
these questions are closely related. For example, some of the homophily
we observed might be caused by women seeking to avoid harassment or
sexism from men {[}38{]}, which would clearly be very concerning.
Additionally, Sheltzer and Smith {[}66{]} concluded that `elite' male
academics (defined as recipients of major honours) have a higher
proportion of male students and postdocs than non-elite male academics.
This finding could contribute to the homophily we observed, and is cause
for concern since the results might reflect discrimination against women
during hiring {[}20{]}, or avoidance by women of elite research groups
(e.g.~due to gender differences in confidence, or a perception that some
groups are sexist). We also found a little evidence that gender
homophily is detrimental to research quality, in that high-impact
journals tended to have weaker homophily (though the relationship was
very noisy). Assuming that papers published in high-impact journals are
of higher average quality (which is contentious; {[}67{]}), our results
provide non-experimental support for the hypothesis that mixed-gender
teams produce better research than single-gender teams {[}42--48{]}.
Another issue is that if many collaborations are between established
researchers, there will be an excess of male-male collaborations in
fields where women in senior positions are rare; some of the observed
homophily might therefore reflect the elevated gender gap among senior
researchers.

On the other hand, homophily might have more benign causes.
Collaboration is often most enjoyable and productive when working with
like-minded people, who might tend to be same-gendered more often than
not. We also suppose that some people consciously choose to
preferentially collaborate with women in order to help close the gender
gap in the workforce; this would create homophily if women adopt this
strategy more often than men. In support of this interpretation, there
is some evidence that women are more likely than men to promote the work
of female colleagues by inviting them to give talks {[}68,69{]}. Given
that many collaborative research projects unfortunately involve a
gendered division of labour {[}61{]}, working with a same-gendered
colleague may provide exposure to new parts of the research process.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{the-dataset}{%
\subsection{The dataset}\label{the-dataset}}

We used the dataset of PubMed author lists from Holman et al. {[}5{]}.
Briefly, that dataset was created by downloading every article indexed
on PubMed and attempting to infer gender from each authors' given name
using computational methods. Each journal was assigned to one of 107
scientific disciplines, using PubMed's journal categorisations in the
interests of objectivity. Because the present study focuses on
co-authorship, all single-author papers were discarded. We also
discarded all papers for which we could not determine the gender of
every author with \({\ge}95\%\) certainty, in order to simplify the
statistical analysis. To mitigate Wahlund effects caused by variation in
the gender ratio of researchers over time (see below), we only kept
papers with publication dates falling in two one-year time periods,
namely 0-1 or 10-11 years prior to the collection date of the PubMed
data (i.e.~20\textsuperscript{th} August 2016). Lastly, we excluded
journals with fewer than 50 suitable papers. Detailed sample size
information is given in S1 Table.

\hypertarget{calculating-alpha-the-coefficient-of-homophily}{%
\subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}}

Following Bergstrom et al. {[}60{]}, we defined the coefficient of
homophily as \(\alpha = p - q\), where \(p\) is the probability that a
randomly-chosen co-author of a \emph{male} author is a man and \(q\) is
the probability that a randomly-chosen co-author of a \emph{female}
author is a man. Like the Wahlund effect, \(\alpha\) is borrowed from
population genetics; for a set of 2-author papers, it is equivalent to
Wright's coefficient of inbreeding {[}70{]}. Mathematical work
illustrates that \(\alpha\) is closely related to alternative
network-based methods for quantifying homophily {[}71{]}.

To estimate \(\alpha\) for a particular subset of the scientific
literature, we estimated \(p\) as the average proportion of men's
co-authors who are men (averaged across all papers with at least one man
author), and \(q\) as the average proportion of women's co-authors who
are men (averaged across all papers with at least one woman author). To
estimate the 95\% confidence intervals on \(\alpha\) for a given set of
\(n\) papers, we sampled \(n\) papers with replacement 1000 times,
estimated \(\alpha\) on each sample, and recorded the 95\% quantiles of
the resulting 1000 estimates.

As well as calculating \(\alpha\) for all authors, we calculated
\(\alpha\) for first or last authors only. \(\alpha\) was again defined
as \(p - q\), but this time \(p\) was estimated as the average
proportion of male co-authors on papers with a male first (or last)
author, and \(q\) was estimated as the average proportion of male
co-authors on papers with female first (or last) authors. We did not
calculate \(\alpha\) for other authorship positions (e.g.~second or
third authors) because this would necessitate culling the dataset to
include only papers with a sufficiently long author list, complicating
interpretation of the results.

We also calculated \(\alpha\) for papers with 2, 3, 4 or \({\ge}5\)
authors, for all journals that had at least 50 suitable papers from
2015-2016 with the specified author list length.

Our test assumes that the expected value of \(\alpha\) is zero if
authors randomly assort, but for small datasets this assumption is not
always true. Essentially, this issue arises because a person cannot be
their own co-author. In a small dataset comprising \(m\) men and \(f\)
women authors, a man can co-author with \(m - 1\) men while a woman can
co-author with \(m\) men. Thus, the null expectation for \(\alpha\) is a
negative number -- potentially a large one if \(m\) and \(f\) are very
small.

To control for the fact that the null expectation for \(\alpha\) is not
zero for small datasets, we devised an adjusted version of the
coefficient of homophily, which we term \(\alpha'\). Every time we
calculated \(\alpha\) for a set of papers, we also determined the
expected value of \(\alpha\) under the null hypothesis that authors
assort randomly with respect to gender. This was accomplished by
randomly permuting authors across papers 1000 times, recalculating
\(\alpha\), and taking the median. We then calculated \(\alpha'\) by
subtracting the null expectation for \(\alpha\) from the observed value.
We also used the null-simulated \(\alpha\) values to calculate a
two-tailed p-value for the observed value of \(\alpha\); the p-value was
defined as the proportion of null simulations for which
\(|\alpha_{null}| > |\alpha_{obs}|\). We applied false discovery rate
(FDR) correction to each set of p-values to account for multiple testing
{[}72{]}.

As expected, \(\alpha'\) was usually almost identical to \(\alpha\) (S8
Fig), but \(\alpha\) was downwardly biased relative to \(\alpha'\) for
the smallest datasets (S9 Fig). Additionally, the correlation between
\(\alpha'\) and sample size was negligible (\(R^2 < 0.01\)), suggesting
that our calculation of \(\alpha'\) effectively removed the dependence
of \(\alpha\) on sample size. We therefore used \(\alpha'\) in all
analyses.

\hypertarget{minimising-the-wahlund-effect-research-discipline-and-time-period}{%
\subsection{Minimising the Wahlund effect: research discipline and time
period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}}

To minimise bias in \(\alpha'\) due to the Wahlund effect, we restricted
each set of papers to a single research specialty to the greatest extent
allowed by our data. Specifically, we only calculated \(\alpha'\) for
individual journals, since papers from the same journal typically focus
on closely related topics. Although some journals, e.g. \emph{PLOS ONE},
publish research from diverse disciplines with very different author
gender ratios {[}5{]}, calculating \(\alpha'\) for these highly
multidisciplinary journals is still useful as a contrast. The difference
in \(\alpha'\) between highly multidisciplinary and more specialised
journals, e.g. \emph{PLOS ONE} versus \emph{PLOS Computational Biology},
gives an estimate of the extent to which multidisciplinarity within
journals inflates \(\alpha'\).

As well as varying between disciplines, the gender ratio of authors has
changed markedly over time {[}5{]}. Because the gender ratio was more
male-biased in the past, \(\alpha'\) would be inflated if we calculated
it for a sample of papers published over a long enough time frame. To
minimise this effect, we only sampled papers from two one-year periods
(namely 2005-6 and 2015-16). The median change per year in \% (fe)male
authors across journals is below 0.5\% {[}5{]}, and so restricting our
dataset to a single year should prevent temporal changes in gender ratio
from noticeably affecting our estimates of \(\alpha'\).

\hypertarget{minimising-the-wahlund-effect-author-country-of-affiliation}{%
\subsection{Minimising the Wahlund effect: author country of
affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}}

A Wahlund effect could arise even if one calculates \(\alpha'\) for a
single discipline and time period, because of variation in the gender
ratio of researchers from different countries. For example, Holman et
al. {[}5{]} found that PubMed-indexed authors based in Serbia are far
more likely to be women than are authors based in Japan. Therefore, a
dataset containing a mix of papers from teams of authors based in these
two countries would contain an excess of same-sex coauthorships, even if
collaboration were random with respect to gender within each country.

To address this issue, we also analysed every combination of journal and
author country of affiliation for which we had enough data (i.e.~50 or
more papers published in 2015-16). For simplicity, we restricted the
dataset to only include papers for which Holman et al. {[}5{]} had
identified the country of affiliation for all authors on the paper, and
all authors shared the same country of affiliation. Restricting the
dataset in this fashion produced enough data to measure \(\alpha'\) for
325 combinations of journal and country (median: 70 papers and 273
authors per combination).

\hypertarget{calculating-standardised-journal-impact-factor}{%
\subsection{Calculating standardised journal impact
factor}\label{calculating-standardised-journal-impact-factor}}

We obtained the 3-year impact factor for each journal from Clarivate
Analytics (formerly ISI). To account for large differences in impact
factor between disciplines, we took the the residuals from a model with
\(log_{10}\) impact factor as the response and the research discipline
of the journal as a random effect. Thus, journals with a positive
standardised impact factor have a higher mean number of citations than
the average for journals in their discipline. We then used linear
regression and Spearman rank correlation to examine the relationship
between \(\alpha'\) and impact factor across journals.

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

Previous authors {[}66,73{]} have hypothesised that senior scientists
preferentially recruit staff and students of the same gender, and/or
that junior researchers preferentially select same-gendered mentors. In
the majority of PubMed-indexed disciplines, authorship conventions mean
that the first-listed author is often an early-career researcher, while
the author listed last is more likely to be a senior researcher leading
a research team {[}74{]}. Assuming that senior researchers are the main
drivers of homophily and that there are enough papers with three or more
authors, we predict that the last author's gender will be the strongest
predictor of the remaining authors' genders (i.e.~the gender of the last
author will be more salient than that of the first author, or any other
authorship position). This is because the first author's gender would
simply be an imperfect correlate of the true causal effect, while the
last author's gender would be the causal effect itself.

To test whether \(\alpha'\) for last authors tends to be higher than
\(\alpha'\) for first authors for any given dataset, we used a linear
mixed model implemented in the \texttt{lme4} and \texttt{lmerTest}
packages for R, with \emph{authorship position} (first or last) as a
fixed factor, and \emph{journal} and \emph{research discipline} as
crossed random effects. The response variable was \(\alpha'\), and we
weighted each observation by the inverse of the standard error from our
estimate of \(\alpha'\), meaning that more accurate measurements of
\(\alpha'\) had more influence on the results. We used a similar model
to test for a difference in \(\alpha'\) between the 2005-6 and the
2015-16 datasets, with two differences: we fit year range as a two-level
fixed factor (instead of authorship position), and we used \(\alpha'\)
estimated for all authors (not first/last authors) as the response
variable.

The relationship between the gender ratio of authors publishing in a
journal and its \(\alpha'\) value appeared nonlinear (S4 Fig). We
therefore fit a generalised additive model with thin plate regression
spline smoothing, implemented using the \texttt{mgcv} package for R.

To model the relationship between \(\alpha'\) and the number of authors
on the paper, we used a meta-regression model implemented in the R
package \texttt{brms} {[}75{]}. The model incorporated the standard
error associated with each estimate of \(\alpha'\), had author number as
a fixed effect, and journal as a random intercept (to control for
repeated measures of each journal). We also fit a random slope of author
number within journal, thereby allowing the response to author number to
vary between journals. We used the default (weak) priors. The full
output of this model can be viewed in the Online Supplementary Material.

\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}{%
\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}}

In most STEMM disciplines, the gender ratio is more skewed among
established researchers relative to early-career researchers, due both
to women leaving STEMM careers at greater rates (the `leaky pipeline'),
and to historical shortages of women studying STEMM subjects at
university (`demographic inertia') {[}1,5{]}. We hypothesised that this
difference in gender ratio between career stages could potentially
create both Wahlund effects and `reverse' Wahlund effects. For example,
imagine that the majority of collaborations in a particular field are
between students and professors, and that the gender ratio differs
between career stages: we would then see an excess of mixed-gender
coauthorships (heterophily, \(\alpha < 0\)), even if gender has no
direct, causal effect on the choice of coauthors. Similarly, a
hypothetical field in which students work only with students, and
professors with professors, would have apparent gender homophily
(\(\alpha > 0\)).

We can think of no tractable method of controlling for this issue using
our dataset, which contains no reliable information on career stage.
Therefore, we instead decided to derive theoretical expectations for
\(\alpha\) when there is a difference in gender ratio across career
stages, in order to determine if and how this effect should alter our
inferences. For simplicity, our calculations assume there are only two
career stages (`early-career' and `established'), though we expect that
the general conclusions would also apply to a multi-tier career ladder.
Under the null model that gender has no causal effect on collaboration,
we calculated \(\alpha\) for various combinations of the four free
parameters in our simple model. These parameters are: the gender ratio
among early-career researchers (\(x\)-axis of Fig 5), the gender ratio
among established researchers (\(y\)-axis of Fig 5), the frequency of
within- versus between career stage collaborator pairs (rows in Fig 5),
and lastly the frequency of within-stage collaborations that are between
two early-career researchers as opposed to two late-career researchers
(columns in Fig 5). When these four parameters are specified, one can
easily calculate the relative frequencies of collaborator pairs that
involve two men, two women, or a man and a woman. In short, if we have
specified the frequency of women at both career stages, as well as the
frequency of the three possible types of collaboration with respect to
career stage (early-early, early-established, and
established-established), then we can calculate the frequency of
collaborators pairs comprising two women, or a woman and a man, and thus
calcualte \(\alpha\) (see the Online Supplementary Material for the
annotated R code).

\hypertarget{data-and-code-availability}{%
\subsection{Data and code
availability}\label{data-and-code-availability}}

The R scripts used to produce all results, figures and tables can be
viewed online at \url{https://lukeholman.github.io/genderHomophily/}.
The input data from Holman et al. {[}5{]} is archived at
\url{https://osf.io/bt9ya/}.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank Devi Stuart-Fox and Dominique Potvin for their very helpful
comments on the manuscript.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Shaw_2012}{}%
1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
inertia from ongoing gender differences in academia. Proceedings of the
Royal Society of London B. 2012;272: 3736--3741.

\leavevmode\hypertarget{ref-Lariviere_2013}{}%
2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
global gender disparities in science. Nature. 2013;504: 211--213.

\leavevmode\hypertarget{ref-West_2013}{}%
3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
gender in scholarly authorship. PLoS ONE. 2013;8: e66212.

\leavevmode\hypertarget{ref-Elsevier_report}{}%
4. Elsevier Report. Gender in the global research landscape.
elseviercom/research-intelligence/resource-library/gender-report. 2017;

\leavevmode\hypertarget{ref-Holman_2018}{}%
5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
long until women are equally represented? PLoS Biology. 2018;16:
e2004956.

\leavevmode\hypertarget{ref-Wutte_2007}{}%
6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.

\leavevmode\hypertarget{ref-Reuben_2014}{}%
7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
careers in science. Proceedings of the National Academy of Sciences.
2014;111: 4403--4408.

\leavevmode\hypertarget{ref-Trower_2002}{}%
8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
underrepresented in the professoriate, and fresh ideas to induce needed
reform. Harvard Magazine. 2002;104: 33--37.

\leavevmode\hypertarget{ref-Umbach_2007}{}%
9. Umbach PD. Gender equity in the academic labor market: An analysis of
academic disciplines. Research in Higher Education. 2007;48: 169--192.

\leavevmode\hypertarget{ref-Hosek_2005}{}%
10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
al. Gender differences in major federal external grant programs. RAND
Corporation. 2005;

\leavevmode\hypertarget{ref-pohlhaus_2011}{}%
11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
differences in application, success, and funding rates for NIH
extramural programs. Academic Medicine. 2011;86: 759.

\leavevmode\hypertarget{ref-Zuckerman_1987}{}%
12. Zuckerman H. Persistence and change in the careers of men and women
scientists and engineers. National Academy Press. 1987; 127--156.

\leavevmode\hypertarget{ref-Rosenfeld_1991}{}%
13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
for the Office of Scientific and Engineering Personnel, National
Research Council. 1991;

\leavevmode\hypertarget{ref-Long_1993}{}%
14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
Sex differences and the effects of productivity. American Sociological
Review. 1993; 703--722.

\leavevmode\hypertarget{ref-Hopkins_2013}{}%
15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
publication patterns by gender, race and ethnicity based on a survey of
a random sample of authors. Scientometrics. 2013;96: 515--534.

\leavevmode\hypertarget{ref-ODorchai_2009}{}%
16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
2009--Statistics and indicators on gender equality in science.
Publications Office of the European Union. 2009;

\leavevmode\hypertarget{ref-Feldt_1986}{}%
17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
MI: Office of Affirmative Action. 1986;

\leavevmode\hypertarget{ref-Stack_2004}{}%
18. Stack S. Gender, children and research productivity. Scientometrics.
2004;45: 891--920.

\leavevmode\hypertarget{ref-Lariviere_2011}{}%
19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
Sex differences in research funding, productivity and impact: an
analysis of Québec university professors. Scientometrics. 2011;87:
483--498.

\leavevmode\hypertarget{ref-Moss_2012}{}%
20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
Science faculty's subtle gender biases favor male students. Proceedings
of the National Academy of Sciences. 2012;109: 16474--16479.

\leavevmode\hypertarget{ref-Knobloch_2013}{}%
21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
gender biases favor male students. Science Communication. 2013;35:
603--625.

\leavevmode\hypertarget{ref-Lee_2005}{}%
22. Lee S, Bozeman B. The impact of research collaboration on scientific
productivity. Social Studies of Science. 2005;35: 673--702.

\leavevmode\hypertarget{ref-Wuchty_2007}{}%
23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
production of knowledge. Science. 2007;316: 1036--1039.

\leavevmode\hypertarget{ref-Abramo_2009}{}%
24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
productivity: is there correlation? Higher Education. 2009;57: 155--171.

\leavevmode\hypertarget{ref-Lariviere_2015}{}%
25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
Collaboration and scientific impact since 1900. Journal of the
Association for Information Science and Technology. 2015;66: 1323--1332.

\leavevmode\hypertarget{ref-Long_1992}{}%
26. Long JS. Measures of sex differences in scientific productivity.
Social Forces. 1992;71: 159--178.

\leavevmode\hypertarget{ref-Bozeman_2011}{}%
27. Bozeman B, Monica G. How do men and women differ in research
collaborations? An analysis of the collaborative motives and strategies
of academic researchers. Research Policy. 2011;40: 1393--1402.

\leavevmode\hypertarget{ref-Abramo_2013}{}%
28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
collaboration. Journal of Informetrics. 2013;7: 811--822.

\leavevmode\hypertarget{ref-Badar_2013}{}%
29. Badar K, Hite JM, Badir YF. Examining the relationship of
co-authorship network centrality and gender on academic research
performance: The case of chemistry researchers in pakistan.
Scientometrics. 2013;94: 755--775.

\leavevmode\hypertarget{ref-Lewison_2001}{}%
30. Lewison G. The quantity and quality of female researchers: A
bibliometric study of Iceland. Scientometrics. 2001;52: 29--43.

\leavevmode\hypertarget{ref-Webster_2001}{}%
31. Webster BM. Polish women in science: A bibliometric analysis of
Polish science and its publications. Research Evaluation. 2001;10:
185--194.

\leavevmode\hypertarget{ref-Bozeman_2004}{}%
32. Bozeman B, Corley E. Scientists' collaboration strategies:
implications for scientific and technical human capital. Research
Policy. 2004;33: 599--616.

\leavevmode\hypertarget{ref-Long_1990}{}%
33. Long JS. The origins of sex differences in science. Social Forces.
1990;68: 1297--1316.

\leavevmode\hypertarget{ref-Fuchs_2001}{}%
34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
scientific organizations in Germany. Minerva. 2001;39: 175--201.

\leavevmode\hypertarget{ref-Reskin_1978}{}%
35. Reskin BF. Scientific productivity, sex, and location in the
institution of science. American Journal of Sociology. 1978;83:
1235--1243.

\leavevmode\hypertarget{ref-Wright_2003}{}%
36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
Catherine M amd Germain, Reed KL. Gender differences in academic
advancement: Patterns, causes, and potential solutions in one U.S.
college of medicine. Social Forces. 2003;68: 1297--1316.

\leavevmode\hypertarget{ref-bleidorn2016age}{}%
37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
J, et al. Age and gender differences in self-esteem -- a cross-cultural
window. Journal of Personality and Social Psychology. 2016;111: 396.

\leavevmode\hypertarget{ref-jagsi_2016}{}%
38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
Sexual harassment and discrimination experiences of academic medical
faculty. JAMA. 2016;315: 2120--2121.

\leavevmode\hypertarget{ref-Martin_2014}{}%
39. Martin JL. Ten simple rules to achieve conference speaker gender
balance. PLoS computational biology. 2014;10: e1003903.

\leavevmode\hypertarget{ref-Tower_2007}{}%
40. Tower G, Julie P, Brenda R. A multidisciplinary study of
gender-based research productivity in the world's best journals. Journal
of Diversity Management. 2007;2: 23--32.

\leavevmode\hypertarget{ref-Jordan_2008}{}%
41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
publication productivity of accounting faculty?. Journal of Applied
Business Research. 2008;24: 77--85.

\leavevmode\hypertarget{ref-Britton_2000}{}%
42. Britton DM. The epistemology of the gendered organization. Gender
and Society. 2000;14: 418--434.

\leavevmode\hypertarget{ref-Reagans_2001}{}%
43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
social capital of corporate R\&D teams. Organization Science. 2001;12:
502--517.

\leavevmode\hypertarget{ref-Hong_2004}{}%
44. Hong L, Page SE. Groups of diverse problem solvers can outperform
groups of high-ability problem solvers. Proceedings of the National
Academy of Sciences. 2004;101: 16385--16389.

\leavevmode\hypertarget{ref-Whittington_2008}{}%
45. Whittington KB, Smith-Doerr L. Women inventors in context:
Disparities in patenting across academia and industry. Gender \&
Society. 2008;22: 194--218.

\leavevmode\hypertarget{ref-Bear_2011}{}%
46. Bear JB, Woolley AW. The role of gender in team collaboration and
performance. Interdisciplinary Science Reviews. 2011;36: 46--153.

\leavevmode\hypertarget{ref-Herrera_2012}{}%
47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
leadership and culture. Global Business and Organizational Excellence.
2012;31: 37--48.

\leavevmode\hypertarget{ref-Campbell_2013}{}%
48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
working groups produce higher quality science. PloS ONE. 2013; e79147.

\leavevmode\hypertarget{ref-Ferber_1980}{}%
49. Ferber MA, Teiman M. Are women economists at a disadvantage in
publishing journal articles? Eastern Economic Journal. 1980;6:
1189--193.

\leavevmode\hypertarget{ref-McDowell_1992}{}%
50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
coauthor: Implications for academic promotion. Economic Inquiry.
1992;30: 68--82.

\leavevmode\hypertarget{ref-Ghiasi_2015}{}%
51. Ghiasi G, Larivière V, Sugimoto CR. On the compliance of women
engineers with a gendered scientific system. PloS ONE. 2015;10:
e0145931.

\leavevmode\hypertarget{ref-Crow_2015}{}%
52. Crow MS, Smykla JO. An examination of author characteristics in
national and regional criminology and criminal justice journals,
2008-2010: Are female scholars changing the nature of publishing in
criminology and criminal justice? American Journal of Criminal Justice.
2015;40: 441--455.

\leavevmode\hypertarget{ref-Fahmy_2017}{}%
53. Fahmy C, Young JT. Gender inequality and knowledge production in
criminology and criminal justice. Journal of Criminal Justice Education.
2017;28: 285--305.

\leavevmode\hypertarget{ref-Zettler_2017}{}%
54. Zettler HR, Stephanie M Cardwell, Jessica M C. The gendering effects
of co-authorship in criminology \& criminal justice research. Criminal
Justice Studies. 2017;30: 30--44.

\leavevmode\hypertarget{ref-Jadidi_2017}{}%
55. Jadidi M, Karimi F, Lietz H, Wagner C. Gender disparities in
science? Dropout, productivity, collaborations and success of male and
female computer scientists. Advances in Complex Systems. 2017; 1750011.

\leavevmode\hypertarget{ref-Teele_2017}{}%
56. Teele DL, Kathleen T. Gender in the journals: Publication patterns
in political science. PS: Political Science \& Politics. 2017;50:
433--447.

\leavevmode\hypertarget{ref-Araujo_2017a}{}%
57. Araújo T, Elsa F. The specific shapes of gender imbalance in
scientific authorships: a network approach. Journal of Informetrics.
2017;11: 88--102.

\leavevmode\hypertarget{ref-Araujo_2017b}{}%
58. Araújo T, Elsa F. Big Missing Data: are scientific memes inherited
differently from gendered authorship? arXiv preprint arXiv. 2017;
1706.05156.

\leavevmode\hypertarget{ref-Wahlund_1928}{}%
59. Wahlund S. Zusammensetzung von populationen und
korrelationserscheinungen vom standpunkt der vererbungslehre aus
betrachtet. Hereditas. 1928;11: 65--106.

\leavevmode\hypertarget{ref-bergstrom_2016}{}%
60. Bergstrom T, Bergstrom C, King M, Jacquet J, West J, Correll S. A
note on measuring gender homophily among scholarly authors. 2016;

\leavevmode\hypertarget{ref-macaluso_2016}{}%
61. Macaluso B, Larivière V, Sugimoto T, Sugimoto CR. Is science built
on the shoulders of women? A study of gender differences in
contributorship. Academic Medicine. 2016;91: 1136--1142.

\leavevmode\hypertarget{ref-Bentley_2003}{}%
62. Bentley JT, Adamson R. Gender differences in the careers of academic
scientists and engineers: A literature review. Special Report. 2003;

\leavevmode\hypertarget{ref-Long_2015}{}%
63. Long MT, Leszczynski A, Thompson KD, Wasan SK, Calderwood AH. Female
authorship in major academic gastroenterology journals: A look over 20
years. Gastrointestinal Endoscopy. 2015;81: 1440--1447.

\leavevmode\hypertarget{ref-Bendels_2018}{}%
64. Bendels MH, Bauer J, Schöffel N, Groneberg DA. The gender gap in
schizophrenia research. Schizophrenia Research. 2018;193: 445--446.

\leavevmode\hypertarget{ref-McKenzie_2017}{}%
65. McKenzie K, Ramonas M, Patlas M, Katz DS. Assessing the gap in
female authorship in the journal emergency radiology: Trends over a
20-year period. Emergency Radiology. 2017;24: 641--644.

\leavevmode\hypertarget{ref-sheltzer_2014}{}%
66. Sheltzer JM, Smith JC. Elite male faculty in the life sciences
employ fewer women. Proceedings of the National Academy of Sciences.
2014;111: 10107--10112.

\leavevmode\hypertarget{ref-Garfield_2006}{}%
67. Garfield E. The history and meaning of the journal impact factor.
JAMA. 2006;295: 90--93.

\leavevmode\hypertarget{ref-nittrouer_2018}{}%
68. Nittrouer CL, Hebl MR, Ashburn-Nardo L, Trump-Steele RC, Lane DM,
Valian V. Gender disparities in colloquium speakers at top universities.
Proceedings of the National Academy of Sciences. 2018;115: 104--108.

\leavevmode\hypertarget{ref-debarre_2018}{}%
69. Débarre F, Rode N, Ugelvig L. Gender equity at scientific events.
Evolution Letters. 2018;in press: doi:10.1002/evl3.49.

\leavevmode\hypertarget{ref-wright_1949}{}%
70. Wright S. The genetical structure of populations. Annals of Human
Genetics. 1949;15: 323--354.

\leavevmode\hypertarget{ref-wang_2016}{}%
71. Wang YS, Erosheva EA. On the relationship between set-based and
network-based measures of gender homophily in scholarly publications.
arXiv preprint arXiv:161009026. 2016;

\leavevmode\hypertarget{ref-Benjamini_1995}{}%
72. Benjamini Y, Hochberg Y. Controlling the false discovery rate: a
practical and powerful approach to multiple testing. Journal of the
Royal Statistical Society: Series B. 1995; 289--300.

\leavevmode\hypertarget{ref-Bonham_2017}{}%
73. Bonham KS, Stefan MI. Women are underrepresented in computational
biology: An analysis of the scholarly literature in biology, computer
science and computational biology. PLoS Computational Biology. 2017;13:
e1005134.

\leavevmode\hypertarget{ref-Wren_2007}{}%
74. Wren JD, Kozak KZ, Johnson KR, Deakyne SJ, Schilling LM, Dellavalle
RP. The write position: A survey of perceived contributions to papers
based on byline position and number of authors. EMBO reports. 2007;8:
988--991.

\leavevmode\hypertarget{ref-burkner_2016}{}%
75. Bürkner P-C. brms: An R package for Bayesian multilevel models using
Stan. Journal of Statistical Software. 2016;80: 1--28.

\newpage

\hypertarget{supporting-information}{%
\section{Supporting information}\label{supporting-information}}

\hypertarget{supplementary-figures}{%
\subsection{Supplementary figures}\label{supplementary-figures}}

\textbf{S1 Fig}: Plot showing the percentage of papers that have 1, 2,
3, 4, or \({\ge}5\) authors for each discipline in the dataset of Holman
et al. (2018). This information can also be found in S3 Data.

\textbf{S2 Fig}: Histogram showing the distribution of differences in
\(\alpha'\) between the 2015-16 and 2005-6 samples, where positive
numbers indicate an increase in \(\alpha'\) with time. The mean is
slightly positive (i.e.~0.004), indicating a mild increase in average
\(\alpha'\) with time.

\textbf{S3 Fig}: Histogram showing the difference between \(\alpha'\)
calculated for first and last authors. Positive values mean that
\(\alpha'\) was higher when calculated for first authors, and negative
values mean \(\alpha'\) was higher when calculated for last authors. The
mean is very slightly higher than zero, indicating that \(\alpha'\)
tends to be higher for first authors.

\textbf{S4 Fig}: There is a weakly positive, non-linear relationship
between the gender ratio of authors publishing in a journal, and the
coefficient of homophily (\(\alpha'\)). Specifically, journals with 50\%
women authors or higher tended to have more same-sex coauthorships than
did journals in which most authors are men. This relationship held
whether \(\alpha'\) was calculated for all authors or first authors
only, but not for last authors only. A negative value on the x-axis
denotes an excess of men authors, a positive value denotes an excess of
women authors, and zero denotes gender parity (i.e.~equal numbers of
male and female authors). The lines were fitted using generalised
additive models with the smoothing parameter \(k\) set to 3.

\textbf{S5 Fig}: Histogram of \(\alpha'\) for 325 unique combinations of
journal and country, using data from August 2015 - August 2016. The
white areas denote combinations for which \(\alpha'\) differs
significantly from zero (p \textless{} 0.05, following false discovery
rate correction).

\textbf{S6 Fig}: Plot showing the 68 combinations of journal and author
country of affiliation for which \(\alpha'\) is significantly higher
than expected.

\textbf{S7 Fig}: Histogram showing the estimated degree to which
\(\alpha'\) is inflated by inter-country differences in author gender
ratio, across the 285 journals for which we had adequate data after
restricting the analysis by country. The average inflation in
\(\alpha'\) is negligible, suggesting that Wahlund effects resulting
from inter-country differences have a negligible effect on our estimates
of gender homophily.

\textbf{S8 Fig}: There is a very strong correlation between the values
of \(\alpha\) and \(\alpha'\) calculated for each journal, though in a
handful of cases the difference is considerable. The deviation between
\(\alpha\) and \(\alpha'\) is greatest for journals for which there is a
small sample size (see S9 Fig).

\textbf{S9 Fig}: For journals for which we recovered a small number of
papers (\textless{}100), the unadjusted metric \(\alpha\) was downwardly
biased. This fits our expectations: because researchers cannot be their
own co-authors, small datasets will tend to produce negative estimates
of \(\alpha\) even if authors assort randomly with respect to gender
(see main text). This suggests that \(\alpha'\) is a better measure of
homophily and heterophily, though the improvement is trivial in large
enough samples.

\hypertarget{supplementary-tables}{%
\subsection{Supplementary tables}\label{supplementary-tables}}

\textbf{S1 Table}: Sample sizes for the two datasets, which comprise
papers published in the timeframes August 2005 - August 2006, and August
2015 - August 2016.

\textbf{S2 Table}: Number of journals showing statistically significant
homophily or heterophily, in two one-year periods. The significance
threshold was p \textless{} 0.05, and p-values were adjusted using
Benjamini-Hochberg false discovery rate correction. Note that the power
of our test is lower for the 2005-2006 data because fewer papers were
recovered per journal: thus, it is not meaningful to compare the \%
significant journals (i.e.~11\% vs 24\%) between the two time periods.

\hypertarget{supplementary-datasets}{%
\subsection{Supplementary datasets}\label{supplementary-datasets}}

\textbf{S1 Data}: This spreadsheet shows the \(\alpha\) values
calculated for each journal, in the 2005 and 2015 samples, and for each
type of author (all authors, first authors, and last authors). The
tables gives the impact factor of each journal, the sample size,
\(\alpha\) and \(\alpha'\) and their 95\% CIs, and the p-value from a
2-tailed test evaluating the null hypothesis that \(\alpha\) is zero
(both raw and FDR-corrected p-values are shown).

\textbf{S2 Data}: This file gives the number and percentage of papers
that have 1, 2, 3, 4, or \({\ge}5\) authors for each \emph{journal} in
the dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that the
sample sizes include papers for which the gender of one or more authors
was not determined by Holman et al.

\textbf{S3 Data}: This file gives the number and percentage of papers
that have 1, 2, 3, 4, or \({\ge}5\) authors for each \emph{discipline}
in the dataset of Holman et al. (2018) \emph{PLOS Biology}. Note that
the sample sizes include papers for which the gender of one or more
authors was not determined by Holman et al.

\textbf{S4 Data}: The table shows the distribution of the \(\alpha'\)
values across journals, split by the research discipline. The gender
ratio column shows the percentage of women authors in the sample used to
calculate \(\alpha'\), across all authorship positions. In the last two
columns, the numbers outside parentheses give the number of journals
that deviate statistically significantly from zero, while the numbers
inside parentheses give the number that remain significant after false
discovery rate correction.

\end{document}